---
title: "The Quintessential Billboard Hit"
author: "Brandon Chan"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This project is currently in the works. It will be finished soon!

My goal in this analysis is to use Spotify's song data on billboard hits from 2000 to 2018 to find what defines the quintessential billboard hit.

```{r read}
billboard <- read.csv("~/SCHOOL/Winter 19/billboard_2000_2018_spotify_lyrics.csv", comment.char="#")
library(tidyverse)
library(lubridate)
```

```{r}
head(billboard)
```

First, I start by cleaning up the data.
  1. Turn the date column into datetime objects
  2. Get rid of songs with unknown data values because these songs are not useful for data analysis.
```{r cleanup}
billboard <- billboard %>% 
  mutate(date = parse_date_time(date, order = "mdy")) %>% 
  filter(broad_genre != "unknown" & energy != "unknown" & liveness != "unknown" & tempo != "unknown" & speechiness != "unknown" & acousticness != "unknown" & instrumentalness != "unknown" & danceability != "unknown" & loudness != "unknown" & valence != "unknown") 

billboard <- billboard[!(is.na(billboard$broad_genre) | billboard$broad_genre==""), ]
```

Next, I want to make a single score that can represent how much of a HIT a song was. I will call it the "HITscore" and it will be calculated by summing the rank and the number of weeks the song spent on the charts.
```{r HITscore}
billboard <- billboard %>%
  mutate(HITscore = rank + weeks)
```

With the HITscore, I can now plot it with much of what defines a song.
```{r data analysis}
sort <- within(billboard, broad_genre <- factor(broad_genre, levels = names(sort(table(broad_genre), decreasing = TRUE))))

ggplot(sort, aes(x = broad_genre, y = HITscore)) + geom_col(aes(fill = year)) + xlab("HIT score") + ylab("Genre") + ggtitle("Biggest HIT Genre")
```
We see here that rap absolutely dominates the charts, especially in recent years. Rock, R&B, and Country are losing their touch, while Pop picks up some steam. What this graph tells us is that for the biggest hit song, its gotta be a rap song.

Now I want to hone in on the optimal song qualities. While some of these elements might not work the best together, we'll assume they do for the purposes of this analysis.
```{r}
head(billboard)
billboard$energy <- as.numeric(paste(billboard$energy))
summary(billboard$energy)
billboard <- billboard %>% 
  #left off right here. making energy categories. divide into 3 percentile categories.
  mutate(energy_category = ifelse(energy > 0.0565 & energy))
ggplot(billboard, aes(x = energy, y = HITscore)) + geom_point()
```

